\chapter*{Acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements}

\noindent
Quiero dedicar esta tesis especialmente a mi \textbf{familia}, por su amor incondicional y su apoyo en cada circunstancia. A mi esposa \textbf{Simonetta}, por todo tu amor y apoyo, porque sé que incluso en los momentos más difíciles siempre estuviste ahí para levantarme; porque este no es un esfuerzo individual y eres el mejor equipo. Gracias por acompañarme en cada aventura sin pensarlo y por ser siempre mi fuente de energía.

A mi mamá \textbf{María de los Ángeles}, por estar siempre pendiente de mi bienestar y por su apoyo constante en cada paso. A mi papá \textbf{Gonzalo}, por tus consejos y respaldo en los momentos clave, y por enseñarme con tu ejemplo a mantenerme firme ante los retos. A ambos, por su apoyo desde mis primeros estudios hasta ahora y por todo su amor que siempre me ha acompañado. A mis hermanos \textbf{Ángela} y \textbf{Gonzalo}, por estar ahí para conversar, escucharme y darme su apoyo sin condiciones. Y a \textbf{Fifa} y \textbf{Gol}, por darme tantas alegrías y recordarme disfrutar cada día. Una mención especial a mi grupo más cercano de amigos: \textbf{Jose}, \textbf{Juliana}, \textbf{Ricardo}, \textbf{Daniela} y \textbf{Laura}, por su amistad, risas y apoyo durante todos estos años.

\vspace{0.5cm}

I want to express my sincere gratitude to my \textbf{supervisors} \textbf{Riccardo Rossi} and \textbf{Joaquín Hernández} for being exceptional advisors, guiding me in my training as a researcher, and always supporting me in reaching my goals. Thank you for helping me grow both professionally and personally.

I am especially thankful to my \textbf{closest friend and collaborator} \textbf{Raul Bravo}, who worked side by side with me during my PhD thesis. It was truly an honor to work together, you made this journey smoother, more enriching, and full of valuable lessons.

I am also grateful to my \textbf{internal collaborators} at \textbf{CIMNE}: \textbf{Carlos Roig}, \textbf{Ruben Zorrilla}, \textbf{Nicolas Sibuet}, and \textbf{Marco Lopez}, for all the academic contributions, teamwork, and constant support throughout this thesis. Working with you was a pleasure. I also want to thank all the amazing people at \textbf{CIMNE} with whom I shared so many good moments during lunch breaks, coffee chats, and different activities: \textbf{Nicolò}, \textbf{Mehdi}, \textbf{Gabriel}, \textbf{Uxue}, \textbf{Andrea Gorgi}, \textbf{Shardool}, \textbf{Mirna}, \textbf{Rodi}, \textbf{Michele}, \textbf{Andrea Montanino}, \textbf{Agustina}, \textbf{Runeal}, \textbf{Ignacio}, \textbf{Carlos}, \textbf{Lu}, \textbf{Junior}, \textbf{Ricardo Tosi}, \textbf{Eduard}, \textbf{Pau}, \textbf{Saman}, \textbf{Polytimi}, \textbf{Henning}, \textbf{Mariano}, \textbf{Ivan}, \textbf{Javad}, and everyone who made the daily work so much better.

I would also like to thank my \textbf{broader collaborators}: \textbf{Stefan Boschert}, \textbf{Rosa M. Badia}, \textbf{Jorge Ejarque}, \textbf{Enrique Quintana}, \textbf{Andrés Tomás}, \textbf{Cristian Catalin}, \textbf{Jorge Vazquez}, \textbf{Karim}, \textbf{Giovanni Stabile}, and all the other members of the \textbf{eFlows4HPC project}, for our work together. These collaborations showed me the value of cross-institutional and cross-industry collaboration, broadened my perspective beyond academia, and strengthened my drive for research that has real impact across domains.

I am deeply grateful to \textbf{Professor Charbel Farhat} for inviting me as part of the \textbf{Fulbright program} and giving me all the tools for a fruitful collaboration. I am also thankful for my direct work with \textbf{Radek Tezaur} and \textbf{Carlos Hernández}, which made this experience even more rewarding. I also thank the whole \textbf{Stanford FRG team}: \textbf{Rodrigo}, \textbf{Jeremy}, \textbf{Faisal}, \textbf{Bruce}, \textbf{Brahmini}, \textbf{Taggart}, \textbf{Christian}, \textbf{Ali}, and the rest of the team, for the enriching discussions during my visit and for welcoming me as one more member of the group.

I truly appreciate the help and support of the \textbf{administrative staff} who guided me along the way: \textbf{Jordi Jiménez}, \textbf{Jazmin Rios}, \textbf{Silvia Aranda}, \textbf{Maria Roses}, \textbf{David Luna}, as well as \textbf{Carinna}, \textbf{Linnea}, and \textbf{Victoria} from \textbf{Fulbright} for their constant dedication and assistance.

I gratefully acknowledge the financial support of the \textbf{FI-SDUR program} (Department of Research and Universities, \textbf{Generalitat de Catalunya}), which funded the majority of my PhD research, including international stays, conferences, and summer schools. I also thank the \textbf{Universitat Politècnica de Catalunya (UPC)} and \textbf{CIMNE} for their institutional and financial support throughout my doctoral studies, including access to research infrastructure and funding for scientific dissemination. In this context, I extend my appreciation to the \textbf{Kratos Multiphysics Group} for their continuous collaboration and for supporting part of my research efforts. I am also thankful to the \textbf{eFlows4HPC European project} for providing access to tools and resources, and for supporting several dissemination and publication efforts related to this thesis. Additionally, I acknowledge the \textbf{X4HPC network} and \textbf{Barcelona Activa} for their support in my technology transfer and entrepreneurship training, through workshops, mentoring, and innovation programs. I further thank the \textbf{Stanford University Department of Aeronautics and Astronautics} for partially funding my research visit and providing a stimulating academic environment.  I am deeply grateful to the \textbf{Fulbright Program} for funding my Visiting Student Researcher stay at Stanford University, which marked a key stage in my academic and professional development.

\textit{This international research stay at Stanford University lasted eight months (October 2024 – June 2025) and meets the requirements for the International Mention of the PhD.}

Lastly, I want to thank all my \textbf{football teammates} from every team I have played with; you made every match unforgettable. And to my big \textbf{family in Mexico}, including all my lifelong friends from my hometown, thank you for always being there. The list would be way too long to include here, but you know who you are. And to everyone else who contributed in any way to this journey, thank you.

